PACE Solver Description: OCMu64,\\ a solver for One-sided Crossing Minimization

### 2024-09-20 tex/paper.tex
Revised remark 6

\section{Definitions}
We use $<$ to compare vertices in $A$ in their fixed ordering. We generalize to weighted graphs
for the proof of \cref{lem:sfp}: a node $u \in B$ is taken to be a function $u: A \to \mathbb
R^{\geq 0}$, where weight 0 represents an absent edge. We write $W_u = \sum_{a \in A}u(a)$ for
the total weight of $u$, and set $\bar u = u/W_u$. Write $N(u) = \supp(u) \subseteq A$ for the
set of neighbors of $u$.

@@

\begin{definition}[Strongly fixed pair]\label{def:sfp}
    We call $u, v \in B$ a \emph{strongly fixed pair} if for every $b \in A$ we have ${\sum_{a
    \leq b} \bar u(a) \geq \sum_{a \leq b} \bar v(a)}$, and at least one of these inequalities
    is strict. Note that this implies $r(u, v) < 0$.
\end{definition}
\begin{lemma}\label{lem:sfp} When $u, v$ is a strongly fixed pair,
    then for any $w: A\to \mathbb R^{\geq 0}$ we have 
    \[
        r(w, \bar v) \leq r(w, \bar u).
    \]
\end{lemma}
\begin{proof}
    Consider the least element $a_0 \in A$ such that $\bar u(a_0) \neq
    \bar v(a_0)$. We must have $\bar u(a_0) > \bar v(a_0)$. Now consider the transformation of
    $\bar u$ which ``moves'' $\delta := \bar u(a_0) - \bar v(a_0)$ of weight from $a_0$ to its
    successor $a_1 \in A$, and call this transformed function $\bar u'$. Then
    \[
        r(w, \bar u') = r(w, \bar u) - \delta w(a_0) - \delta w(a_1) \leq r(w, \bar u).
    \]
    Since $\bar v$ is obtained from $\bar u$ by a sequence of such transformations,
    the inequality follows.
\end{proof}
\begin{lemma}\label{lem:sfimplf}
    If $(u, v)$ is strongly fixed, then $u \b v$.
\end{lemma}
\begin{proof}
    Suppose towards a contradiction that $v < x_0 < \cdots < x_k < u$ is part of an optimal
    solution. Write $X = \sum_i x_i$ for the combined function. Then by assumption $r(X, u)
    \leq 0$, and therefore $r(X, v) = W_v r(X, \bar v) \leq W_v r(X, \bar u) = W_v / W_u \cdot
    r(X, u) \leq 0$. But then $c(X, u, v) < c(X, v, u)\leq c(v, X, u)$, which
    contradicts $(v,X,u)$ being optimal.
\end{proof}

@@

\begin{remark}\label{rem:strongest}
    By inspecting only $u, v$, it is not possible to do better than \cref{lem:sfimplf}. If $u,
    v$ with $\bar u \neq \bar v$ are not strongly fixed, and we can assume the existence of any
    additional nodes in $A, B$, then we can conjure a situation where $v \b u$.

    Suppose that $b \in A$ be some element such that $U := \sum_{a \leq b} \bar u(a) < \sum_{a \leq
    b} \bar v(a) =: V$, and take $M \in A$ to be some element between $b$ and its succesor. Also,
    take $L < \min(N(v))$ and $R > \max(N(u))$. Now consider a new element $X \in B$, which
    connects only to $L, M, R$. We claim that by picking the right weights for $X$, we can
    obtain $c(v, X, u) < \min(c(X, u, v), c(u, v, X))$, which establishes that in the presence
    of $X$, $v \b u$.

    We have that
    \[
        c(X, u, v) - c(v, X, u) = r(u, v) + W_v(2VX(M) + X(R) - X(L) - X(M))
    \]
    and
    \[
        c(u, v, X) - c(v, X, u) = r(u, v) + W_u(X(L) + (1-U)X(M) - X(R) - UX(M)),
    \]
    and if we can choose $X(L), X(M), X(R)$ so that both these values are positive, we are
    done. We can always scale these values up, so we can ignore the $r(u, v)$ term in both
    equations; we can also ignore $W_u, W_v$. Because we can choose $X(R), X(L)$ as we please,
    it suffices if the sum of the remaining pairs of expressions, which is
    \[
        (2V - 1)X(M) - (2U - 1)X(M),
    \]
    is positive. But indeed it is, since by assumption $U < V$.

    Suppose that $u, v \in B$ with $\bar u \neq \bar v$ are not strongly fixed, and let $b \in
    A$ be some element such
    that $\sum_{a \leq b} \bar u(a) > \sum_{a \leq b} \bar v(a)$. By taking $X: A \to \mathbb
    R^{\geq 0}$ some function which assigns very high
    weight to $b$, we can obtain $c(v, X, u) < \min(c(X, u, v), c(u, v, X))$, showing that
    \cref{lem:sfp} is optimal.
\end{remark}
Although such a function $X$ may exist in theory, it does
not have to exist in the actual set $B$, motivating the following definition
that generalizes RRLO2 of \cite{dujmovic_2008}.

>>>>

\section{Definitions}
We use $<$ to compare vertices in $A$ in their fixed ordering. We generalize to weighted graphs
for the proof of \cref{lem:sfp}: a node $u \in B$ is taken to be a function $u: A \to \mathbb
R^{\geq 0}$, where weight 0 represents an absent edge. Write $N(u) = \supp(u) \subseteq A$ for the
set of neighbors of $u$. We write $W_u = \sum_{a \in A}u(a)$ for the total weight of $u$, and
set $\bar u = u/W_u$. We think of $\bar u$ as a probability distribution, and also consider the
cumulative distribution function $F_{\bar u}(b) = \sum_{a \leq b} \bar u(a)$.

@@

\begin{definition}[Strongly fixed pair]
    We call $u, v \in B$ a \emph{strongly fixed pair} if for every $b \in A$ we have $F_{\bar
    u}(b) \geq F_{\bar v}(b)$, and at least one of these inequalities
    is strict. Note that this implies $r(u, v) < 0$.
\end{definition}
\begin{lemma}\label{lem:sfp} When $u, v$ is a strongly fixed pair,
    then for any $w: A\to \mathbb R^{\geq 0}$ we have 
    \[
        r(w, \bar v) \leq r(w, \bar u).
    \]
\end{lemma}
\begin{proof}
    Consider the least element $a_0 \in A$ such that $\bar u(a_0) \neq
    \bar v(a_0)$. We must have $\bar u(a_0) > \bar v(a_0)$. Now consider the transformation of
    $\bar u$ which ``moves'' $\delta := \bar u(a_0) - \bar v(a_0)$ of weight from $a_0$ to its
    successor $a_1 \in A$, and call this transformed function $\bar u'$. Then
    \[
        r(w, \bar u') = r(w, \bar u) - \delta w(a_0) - \delta w(a_1) \leq r(w, \bar u).
    \]
    Since $\bar v$ is obtained from $\bar u$ by a sequence of such transformations,
    the inequality follows.
\end{proof}
\begin{lemma}\label{lem:sfimplf}
    If $(u, v)$ is strongly fixed, then $u \b v$.
\end{lemma}
\begin{proof}
    Suppose towards a contradiction that $v < x_0 < \cdots < x_k < u$ is part of an optimal
    solution. Write $X = \sum_i x_i$ for the combined function. Then by assumption $r(X, u)
    \leq 0$, and therefore $r(X, v) = W_v r(X, \bar v) \leq W_v r(X, \bar u) = W_v / W_u \cdot
    r(X, u) \leq 0$. But then $c(X, u, v) < c(X, v, u)\leq c(v, X, u)$, which
    contradicts $(v,X,u)$ being optimal.
\end{proof}

@@

\begin{remark}\label{rem:strongest}
    Suppose that $u, v$ with $\bar u \neq \bar v$ are not strongly fixed, and let $b \in A$ be
    such that $F_{\bar u}(b) < F_{\bar v}(b)$. Furthermore suppose that there are nodes $l, m,
    r \in A$, with $m$ sitting between $b$ and its successor in $A$, and $l < (\supp(u) \cup
    \supp(v)) < r$. Then consider a node $X \in B$ which is connected to $l, m, r$. By choosing
    the weights of $X$ appropriately, we can let both $r(X, v) = W_v r(X, \bar v)$ and $r(u, X)
    = W_u r(\bar u, X)$ grow arbitrarily large. Since we can scale up the weights of $X$, it
    suffices to make $r(X, \bar v)$ and $r(\bar u, X)$ simultaneously positive.

    Now $r(X, \bar v) = (2F_{\bar v}(b) - 1)X(m) + (X(r) - X(l))$, and $r(\bar u, X) = (1 -
    2F_{\bar u}(b))X(m) - (X(r) - X(l))$. By choosing $X(r), X(l)$ appropriately, we can make
    these two values equal, so it suffices if their \emph{average} is positive; and indeed 
    \[
        (r(X, \bar v) + r(\bar u, X))/2 = X(m)(F_{\bar v}(b) - F_{\bar u}(b)) > 0.
    \]
    Therefore, the right choice of weights of $X$ creates a situation where $c(v, X, u) <
    \min(c(X, u, v), c(u, v, X))$. This demonstrates that if you want to show $u \b v$, and $u,
    v$ are not strongly fixed, you must consider other features of the graph than $u, v$ and
    their weights: \cref{lem:sfp} is optimal. Note: this also applies in the unweighted case,
    by taking $l, m, r$ to be sets of nodes rather than single nodes.
\end{remark}
Although such a node $X$ may exist in theory, it does
not have to exist in the actual set $B$, motivating the following definition
that generalizes RRLO2 of \cite{dujmovic_2008}.
